---
title: "03 Add a Base Map to a Static Map"
output: html_notebook
---

Using the `maptiles` package to add an OS Maps API base map to a plot of an `sf`
`data.frame`.

## `maptiles`

[`maptiles`](https://cran.r-project.org/package=maptiles) allows pre-rendered
web mapping tiles to be easily added to a data.frame plot as a base map.

## OS Data Hub

The OS Data Hub is the new data portal to access Ordnance Survey (OS) data. Data
can be downloaded through a user interface or programmatically via a series of
APIs enabling direct integration with software applications.

The OS Data Hub [Public Sector
Plan](https://www.ordnancesurvey.co.uk/business-government/public-sector-geospatial-agreement/data-hub-for-public-sector)
provides [PSGA
members](https://www.ordnancesurvey.co.uk/business-government/sectors/public-sector)
with unlimited access to OS OpenData and OS Premium data available under the
PSGA via API or download.

---

```{r}
library(sf)
library(tmap)
library(maptiles)
```

## Create an `sf` object from GeoPackage (GPKG)

```{r}
# Create an sf data.frame object from the Greenspace file
osogs <- st_read('../../data/ordnance-survey/os-open-greenspace-gb.gpkg',
                 layer = 'greenspace_site')
```

## Spatially subset data frame

Using coordinates to spatially subset by bounding box (BBOX).

```{r}
# Subset the data by intersection with BBOX

# Greater London bounding box
bbox <- st_bbox(c(xmin = 503568.1996, 
                  xmax = 561957.4962, 
                  ymin = 155850.7975, 
                  ymax = 200933.9026), 
                crs = st_crs(27700))

# Convert into a 'geometry' sfc = simple feature collection
bbox <- st_as_sfc(bbox)

# Spatial subsetting/filtering rows - does polygon touch any part of bounding box?
osogs <- osogs[bbox, ]
```

```{r}
# Row count
nrow(osogs)
```

```{r}
# Reproject to Web Mercator (EPSG: 3857)
# Prepare for basemap tiles
osogs <- st_transform(osogs, crs = 3857)
```

## Plot `sf` data frame

```{r}
# Demonstrate basic plotting using the built-in functions
plot(st_geometry(osogs), col = '#00cd6c', border = NA)
```

```{r}
# Basic plot using tmap
p <- tm_shape(osogs) +
  tm_fill(col = '#00cd6c') +
  tm_layout(title = 'Spatial Distribution of OS Open Greenspace in Greater London Authority',
            frame = FALSE)

p
```

## Add OS Maps API base map to `tmap` plot

```{r}
# OS Maps API layer
# Example uses Light Style in Web Mercator (EPSG:3857) projection
layer <- 'Light_3857'

# OS Data Hub project API key
key <- '7UTXMMWsGjjIzcBmLdAGnMO6WEAQi9Ng'

# OS Data Hub base path - https://api.os.uk
# OS Maps API ZXY endpoint path - /maps/raster/v1/zxy/
url <- paste0('https://api.os.uk/maps/raster/v1/zxy/', layer,
              '/{z}/{x}/{y}.png?key=XXXXXX')  # Note the placeholder key
```

```{r}
# Define the tile server parameters for maptiles
osmaps <- list(src = 'OS Maps',
               q = url,
               sub = '',
               cit = paste0('Contains OS data © Crown copyright and database rights ', 
                            format(Sys.Date(),"%Y"), '.'))
```

```{r}
# Download tiles and compose base map raster
tile_maps <- get_tiles(x = st_bbox(osogs),
                       provider = osmaps,
                       zoom = 10,  # If omitted, will search for 'best' zoom
                       crop = TRUE,
                       cachedir = tempdir(),  # Where to store the download
                       apikey = key,
                       verbose = FALSE)
```

```{r, fig.width = 10, fig.height = 10}
# Add basemap to the tmap
final_plot <- tm_shape(tile_maps,
                       bbox = st_bbox(tile_maps)) +
                tm_rgb() +
                tm_credits(osmaps$cit,
                           position = c("LEFT", "BOTTOM")) +   # add citation info to the map
                p          # Layer greenspace plot on top of basemap

final_plot
```
