---
title: "05 Display an Interactive Map"
output: html_notebook
---

Using the `tmap` package to display a basic interactive map using the OS Data
Hub OS Maps API.

## `tmap`

The [`tmap`](https://r-tmap.github.io/tmap/) R package is providing a wrapper
around the JavaScript web mapping library [Leaflet.js](https://leafletjs.com/).
It allows the visualisation of geospatial data in R on an interactive leaflet
map. The R package [`leaflet`](https://rstudio.github.io/leaflet/) is an
alternative tool for generative interactive maps.

## OS Data Hub

The OS Data Hub is the new data portal to access Ordnance Survey (OS) data. Data
can be downloaded through a user interact or programmatically via a series of
APIs enabling direct integration with software applications.

The OS Data Hub [Public Sector
Plan](https://www.ordnancesurvey.co.uk/business-government/public-sector-geospatial-agreement/data-hub-for-public-sector)
provides [PSGA
members](https://www.ordnancesurvey.co.uk/business-government/sectors/public-sector)
with unlimited access to OS OpenData and OS Premium data available under the
PSGA via API or download.

## OS Maps API

The [OS Maps API](https://osdatahub.os.uk/docs/wmts/overview) is a mapping API
available through the OS Data Hub. It allows OS base mapping to be embedded into
mapping applications including a [tiled web
map](https://en.wikipedia.org/wiki/Tiled_web_map) or slippy map in R. The API
serves pre-rendered raster tiles in two coordinate reference systems (CRS),
British National Grid ([EPSG:27700](https://epsg.io/27700)) and Web Mercator
([EPSG:3857](https://epsg.io/3857)) which geospatial data can be overlayed onto
of. The raster tiles are also available in a range of styles to support
different geodata visualisation use cases.

Use of the OS Maps API for both OpenData and Premium data map views is free to
the Public Sector but requests still require an [API
key](https://en.wikipedia.org/wiki/Application_programming_interface_key) for
authenitcation. If you do not have access to the OS Data Hub under the Public
Sector plan then you can still use the OS Maps API for free via the [OS OpenData
plan](https://osdatahub.os.uk/plans).

## ZXY vs WMTS

The OS Maps API offers two different resources, ZXY and Web Map Tile Service
(WMTS). Both resources are supported by `R` but the code required to integrate
a ZXY resource is less verbose and preferred.

With the ZXY resource the Z, X and Y provide the indices of a pre-rendered
mapping tile within a request. Leaflet translates the extent of the map for a
given zoom level (Z) into a series of API requests that will download the tiles
required to fill the extent.

---

```{r}
library(tmap)
```

## OS Maps API ZXY resource

```{r}
# OS Maps API layer
# Example uses Outdoor Style in Web Mercator (EPSG:3857) projection
layer <- 'Outdoor_3857'

# Other options: 'Road_27700', 'Road_3857', 'Outdoor_27700', 'Outdoor_3857', Light_3857, 'Light_27700', 'Leisure_27700'

# OS Data Hub project API key
key <- '7UTXMMWsGjjIzcBmLdAGnMO6WEAQi9Ng'
```

```{r}
# Define the tile server parameters for the basemap
url <- paste0('https://api.os.uk/maps/raster/v1/zxy/', layer,
              '/{z}/{x}/{y}.png?key=', key)
```

## Create custom map

```{r, fig.width = 10, fig.height = 10}
# Change plotting mode from static to interactive
tmap_mode('view')

# Display an interactive map
tm_tiles(server = url)
```

```{r}
# Change back to the static plotting mode
tmap_mode('plot')
```

---

## Alternative method using `leaflet`

Demonstration using a different `R` package.

```{r}
library(leaflet)
library(leafem)
library(dplyr)
```

```{r}
# OS Maps API layer
# Example uses Light Style in Web Mercator (EPSG:3857) projecdtion
layer <- 'Light_3857'

# OS Data Hub project API key
key <- '7UTXMMWsGjjIzcBmLdAGnMO6WEAQi9Ng'
```

```{r}
# Define the tile server parameters for the basemap
url <- paste0('https://api.os.uk/maps/raster/v1/zxy/', layer,
              '/{z}/{x}/{y}.png?key=', key)
```

```{r}
# OS Logo image
logo_url <- 'https://raw.githubusercontent.com/OrdnanceSurvey/os-api-branding/master/img/os-logo-maps.svg'
```

## Create custom map

```{r}
# m is the map object all other layers will be added to.
m <- leaflet() %>%
  setView(lng = -0.105, lat = 51.507, zoom = 12)
```

```{r, fig.width = 10, fig.height = 10}
m <- m %>% 
  addTiles(urlTemplate = url,
           attribution = paste0('Contains OS data © Crown copyright and database right ', 
                                format(Sys.Date(), "%Y")),
           group = 'OS Maps') %>%
  addLogo(logo_url, url = 'https://www.ordnancesurvey.co.uk/')

# Return map
m
```
