---
title: "06 Add an Overlay to an Interactive Map"
output: html_notebook
---

Using the `sf` and `tmap` packages to  overlay a spatial object ontop of a
leaflet map with an OS Maps API base map.

---

## `tmap`

The [`tmap`](https://r-tmap.github.io/tmap/) R package is providing a wrapper
around the JavaScript web mapping library [Leaflet.js](https://leafletjs.com/).
It allows the visualisation of geospatial data in R on an interactive leaflet
map.

```{r}
library(sf)
library(tmap)
```

## Create `sf` object from GeoPackage (GPKG)

```{r}
# Create an sf data.frame object from the Greenspace file
osogs <- st_read('../../data/ordnance-survey/os-open-greenspace-gb.gpkg',
                 layer = 'greenspace_site')
```

## Spatially subset data frame

Using coordinate-based indexer to spatially subset by bounding box (BBOX).

```{r}
# Subset the data by intersection with BBOX

# Greater London bounding box
bbox <- st_bbox(c(xmin = 503568.1996, 
                  xmax = 561957.4962, 
                  ymin = 155850.7975, 
                  ymax = 200933.9026), 
                crs = st_crs(27700))

# Convert into a 'geometry'
bbox <- st_as_sfc(bbox)

# Spatial subsetting/filtering
osogs_filtered <- osogs[bbox, ]
```

## OS Maps API ZXY resource

```{r}
# Reproject to Web Mercator (EPSG: 3857)
osogs_filtered <- st_transform(osogs_filtered, crs = 3857)
```

```{r}
# Basic plot using tmap
p <- tm_shape(osogs_filtered) +
  tm_fill(col = '#00cd6c') +
  tm_layout(frame = FALSE)
```

```{r}
# OS Maps API layer
# Example uses Light Style in Web Mercator (EPSG:3857) projection
layer <- 'Light_3857'

# OS Data Hub project API key
key <- '7UTXMMWsGjjIzcBmLdAGnMO6WEAQi9Ng'

# Define the tile server parameters for the basemap
url <- paste0('https://api.os.uk/maps/raster/v1/zxy/', layer,
              '/{z}/{x}/{y}.png?key=', key)
```

## Create custom interactive map

```{r, fig.width = 10, fig.height = 10}
# Change plotting mode from static to interactive
tmap_mode('view')

# Combine the features and base map
final_plot <- p +
                tm_basemap(server = url)

final_plot
```

```{r}
# Change back to the static plotting mode
tmap_mode('plot')
```

## Alternative method using `leaflet`

Demonstration using a different `R` package.

```{r}
library(leaflet)
library(dplyr)
```

## Create custom map

```{r}
# m is the map object all other layers will be added to.
m <- leaflet() %>%
  setView(lng = -0.105, lat = 51.507, zoom = 12)
```

```{r, fig.width = 10, fig.height = 10}
m <- m %>% 
  addTiles(urlTemplate = url,
           attribution = paste0('Contains OS data © Crown copyright and database right ', 
                                format(Sys.Date(), "%Y")),
           group = 'OS Maps') %>%
  addPolygons(data = st_transform(osogs_filtered, 4326),
              color = '#00cd6c')

# Return map
m
```
