---
title: "01 OS GeoDataViz Toolkit"
output: html_notebook
---

Using the OS GeoDataViz Toolkit to assist with styling of location data
insights.

## OS GeoDataViz Toolkit

The OS [GeoDataViz (GDV)
Toolkit](https://github.com/OrdnanceSurvey/GeoDataViz-Toolkit) provides a set of
resources to help you communicate your data effectively through the design of
compelling visuals.

The toolkit provides a series of colour palettes which provide a great starting
point for thematic map design. The palettes can be programmatically accessed via
JSON copy in the repo.

## Qualitative Styling

When styling qualitative (categorical) data we need to be mindful of the number
of distinct classes. The GDV qualitative colours supports up to eight classes.
Beyond approximately 10 classes it becomes very difficult to design a palette
that preserves the separation between colours allowing clear class
identification.

---

```{r}
library(sf)
library(tmap)
library(httr)
library(jsonlite)
```

## Create `sf` data frame from GeoPackage (GPKG)

Combine the file reading and spatial subsetting by bounding box.

```{r}
# Creating bounding box
bbox <- st_bbox(c(xmin = 503568.1996, 
                  xmax = 561957.4962, 
                  ymin = 155850.7975, 
                  ymax = 200933.9026), 
                crs = st_crs(27700))

# Convert into a 'geometry'
bbox <- st_as_sfc(bbox)

# Create an sf data.frame object from the Greenspace file
# Can do subset using bounding box at same time as read in
osogs <- st_read('../../data/ordnance-survey/os-open-greenspace-gb.gpkg',
                 layer = 'greenspace_site',
                 wkt_filter = st_as_text(bbox))  # Subset the data to Greater London
```

## Subset data frame by function

Reduce the number of function categories to support qualitative styling colour
palette integration.

```{r}
# Subset data frame by function, excluding 'Play Space' and 'Tennis Court'
osogs_filtered <- osogs[!osogs$function. %in% c('Play Space', 'Tennis Court'), ]
```

```{r}
# List unique function values
unique(osogs_filtered$function.)
```

```{r}
# Count number of unique function values
length(unique(osogs_filtered$function.))
```

## OS Maps API ZXY resource

```{r}
# OS Maps API layer
# Example uses Light Style in Web Mercator (EPSG:3857) projection
layer <- 'Light_3857'

# OS Data Hub project API key
key <- '7UTXMMWsGjjIzcBmLdAGnMO6WEAQi9Ng'
```

```{r}
# Define the tile server parameters
url <- paste0('https://api.os.uk/maps/raster/v1/zxy/', layer,
              '/{z}/{x}/{y}.png?key=', key)
```

## OS GeoDataViz colour palettes

The OS [GeoDataViz
toolkit](https://github.com/OrdnanceSurvey/GeoDataViz-Toolkit/tree/master/Colours)
provides qualitative, sequential, and diverging colour palettes to support GDV
applications.

```{r}
# GDV colour palettes JSON file
gdv <- 'https://raw.githubusercontent.com/OrdnanceSurvey/GeoDataViz-Toolkit/master/Colours/GDV-colour-palettes-v0.7.json'

# Make HTTP GET request and decode JSON
gdv_json <- GET(url = gdv)
gdv_json <- fromJSON(content(gdv_json))

# Get colour hex values
gdv_qual <- unlist(gdv_json$qualitative$lookup)

# Show an example of the colours
(image(1:length(gdv_qual), 1, as.matrix(1:length(gdv_qual)), 
      col = gdv_qual,
      xlab="", ylab = "", xaxt = "n", yaxt = "n", bty = "n"))
```

## Plot data frame ontop of interactive map

Employ categorical styling via OS GeoDataViz (GDV) qualitative colour palette.

```{r}
# Generate a basic plot of the features
p <- tm_shape(osogs_filtered) +
      tm_polygons('function.', 
                  pal = gdv_qual,
                  popup.vars = FALSE)
```

```{r, fig.width = 10, fig.height = 10}
# Change plotting mode from static to interactive
tmap_mode('view')

# Combine the features and base map
final_plot <- p +
                tm_basemap(server = url)

final_plot
```

```{r}
# Change back to the static plotting mode
tmap_mode('plot')
```
