---
title: "03 GDV Classification Schemes"
output: html_notebook
---

Choropleth map classification scheme examples.

```{r}
library(sf)
library(classInt)
library(ggplot2)
```

## Create data frame from GeoPackage (GPKG)

```{r}
# Create an `sf` data frame from a GeoPackage (GPKG)
osogs_pop_est_lad_merged <- st_read('../../data/lad-os-open-greenspace-area-per-head.gpkg')
```

```{r}
# List names of attributes
names(osogs_pop_est_lad_merged)
```

```{r}
# Return head of data frame
head(osogs_pop_est_lad_merged)
```

## Data Classification

Categorising data into groups for plotting can be seen as a classification
problem. The method chosen should take into account the goal of the map (e.g.
highlighting outliers vs. depicting the distribution of values). To help
decide on the breaks to use please see these helpful links 
[here](https://censusgis.wordpress.com/students/lesson-5-visualisation-cartographic-practice/)
and [here](https://gisgeography.com/choropleth-maps-data-classification/).

The R package [`classInt`](https://r-spatial.github.io/classInt/index.html)
which is used by `tmap`, can be used for choropleth map classification.
Classification schemes can be used via the `style` keyword argument in `tm_*`
calls.

The content below looks at four commonly used classification schemes:

1. Equal Interval
2. Jenks (Natural Breaks)
3. Quantiles
4. Standard Deviation

### Equal Interval

The boundaries between classes are at regular intervals. It is simple and easy
to interpret. However, it considers the extreme values of the distribution
and, in some cases, this can result in one or more classes being sparse.

```{r}
ei5 <- classIntervals(osogs_pop_est_lad_merged$area2population, 
                      n = 5, style = 'equal')

ei5
```

### Jenks Natural Breaks

The classes are defined according to natural groupings of data values. It
minimises differences between points within a class and maximises between
classes differences (clusters data into similar groups).

```{r}
fj5 <- classIntervals(osogs_pop_est_lad_merged$area2population, 
                      n = 5, style = 'jenks')

fj5
```

### Quantiles

Splits the data into classes so there are the same number of data values in each
class. Numeric size of each class is rigidly imposed and can lead to misleading
interpretation. The placement of boundaries may assign almost identical
attributes to adjacent classes, or features with quite widely different values
in the same class.

```{r}
q5 <- classIntervals(osogs_pop_est_lad_merged$area2population, 
                     n = 5, style = 'quantile')

q5
```

### Standard Deviation

Shows the distance of the observation from the mean and then generates class
breaks in standard deviation measures above and below it. Can be good for
identifying outliers.

```{r}
msd <- classIntervals(osogs_pop_est_lad_merged$area2population, 
                      n = 5, style = 'sd')

msd
```

## Compare classification schemes

```{r}
ggplot(osogs_pop_est_lad_merged, aes(x = area2population)) +
  geom_density(color = 'darkblue', fill = 'lightblue') + 
  geom_rug(color = 'red') +
  geom_vline(xintercept = ei5$brks, color = 'red') +
  ggtitle('Equal Interval')
```

```{r}
ggplot(osogs_pop_est_lad_merged, aes(x = area2population)) +
  geom_density(color = 'darkblue', fill = 'lightblue') + 
  geom_rug(color = 'red') +
  geom_vline(xintercept = fj5$brks, color = 'red') +
  ggtitle('Jenks')
```

```{r}
ggplot(osogs_pop_est_lad_merged, aes(x = area2population)) +
  geom_density(color = 'darkblue', fill = 'lightblue') + 
  geom_rug(color = 'red') +
  geom_vline(xintercept = q5$brks, color = 'red') +
  ggtitle('Quantiles')
```

```{r}
ggplot(osogs_pop_est_lad_merged, aes(x = area2population)) +
  geom_density(color = 'darkblue', fill = 'lightblue') + 
  geom_rug(color = 'red') +
  geom_vline(xintercept = msd$brks, color = 'red') +
  ggtitle('Std. Deviation')
```
